LME-MPM applied to quasi-brittle fracture.

### 2020-05-29 2020_EFM_MPM_Eigensoftening.tex
Cambios 29-5-2020

\newglossaryentry{nu}
{
  name=$\nu$,
  description={Poisson ratio}
}
\begin{itemize}
\item 
\end{itemize}

@@

\subsection{Fracture modelling approach}
\label{sec:2.3}
Within the context of \acrshort{mpm} formulation, fracture can be
modelled by failing particles according to a suitable failure
criterion. When material points are failed, they are assumed to have
null stress tensor. To reproduce this behaviour in the present
research, the eigensoftening algorithm is introduced in the 
\acrshort{mpm} framework as an alternative approach to the decohesion
model \cite{Zhenmao_2005,Schreyer_2002}. The eigensoftening concept
was originally developed by Navas {\it et al.}
(2017)\cite{Navas_2017_ES} as an extension for quasi-brittle materials
of the eigenerosion proposed by Pandolfi \& Ortiz
(2012)\cite{Pandolfi_2012} for fracture of brittle materials. \DM{ XXXX I would modify the following  sentence as "The evaluation of both approaches \cite{Navas_2017_ES} show that the
eigenerosion algorithm significantly overestimates the tensile  
stress and the strain peaks, meanwhile eigensoftening captures the forces and crack
patterns accurately XXXXX}. A
comparison between both in \cite{Navas_2017_ES} shows that the
eigenerosion algorithm significantly overestimates the tensile  
stress and the strain peaks, meanwhile eigensoftening captures the forces and crack
patterns accurately. On the other hand, eigensoftening algorithm agrees
very well with experimental results in all the aspects \DM{which aspect?} . Furthermore,
the accuracy of this algorithm has been proven for complex fracture
patterns such as the ones obtained with fiber reinforces concrete (FRC),
\cite{Navas_2018_ES}.\\

The key idea behind the eigenerosion algorithm is the computation of the
energy-release rate attendant to the failure of material point $p$,
\begin{align}
  \label{eq:energy-release-EE}
&G_p^{k+1} = \frac{C_{\epsilon}}{m_p^{k+1}}  \sum_{x_q^{k+1} \in
  B_{\epsilon}(x_p^{k+1})} m_q W_q^{k+1}\\
  \label{eq:mass-EE}
&m_p^{k+1} =  \sum_{x_q^{k+1} \in
  B_{\epsilon}(x_p^{k+1})} m_q  
\end{align}
where $B_{\epsilon}(x_p^{k+1})$ is a $^n$dimensional sphere of radius $\epsilon$
centered at $x_p^{k+1}$. The particles which lie under this sphere are
known as the $\epsilon$-neighborhood of the material point (see
\cite{Pandolfi_2012}). This concept is conveniently sketched in
Figure \ref{fig:Failed-particles}.
%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.5\textwidth]{Figure-Particle-failed}
  \caption{Scheme of a linear cohesive law, where the shaded area is
    $G_f$, $f_t$ is the tensile strength, and $w_c$ is the critical
    opening displacement.}
  \label{fig:Failed-particles}
\end{figure}
%%%%%%%%%%%
Other parameters are the mass of the neighborhood $m_p^{k+1}$, the
current free-energy density per unit mass  $W_q^{k+1}$ and the
normalizing constant $C_{\epsilon}$, \PNA{which also defines the $\epsilon$-neighborhood size, being $\epsilon = C_\epsilon h$.}

@@

reached. Once the crack, $w_n$, reaches the value of the
critical crack opening, $w_c$, a stress-free crack is
attained. \DM{ see the following sentence: comma missing? Clarified--->} For intermediate values a damage value between zero and one represents the extension to which the material has failed. \DM{ hasta acá }

@@

initialization was adopted. Particularly the maximum principal stress
theory for brittle fracture was considered  \DM{ delete by the authors}  by authors
\cite{Navas_2017_ES}. With that purpose \DM{ which purpose?}, the variation of the averaged
strain energy density in the $\epsilon$-neighborhood of the material point
$\vec{x}_p^{k+1}$ can be obtained as

@@

where $m_p$ is the total mass of the $\epsilon$-neighborhood, defined similarly  \DM{ Pondría:  defined similarly TO  the eigenerosion method} than in the eigenerosion methodology (Eq.~\eqref{eq:mass-EE}).
%\begin{equation}
%  \label{eq:averaged-mass}
%  m_p = \sum_{x_q^{k+1} \in B_{\epsilon}(x_p^{k+1})} m_q.
%\end{equation}
This definition \DM{ dos veces definition en la misma oración, podes poner directamente: The equivalent critical stress leads to a definition of the averaged strain energy ... } of the equivalent critical stress leads to a
definition of the averaged strain energy in terms of the averaged

@@

\subsection{$\epsilon$-neighborhood reconstruction : A node-linked method}
\label{sec:epsil-neighb-reconst}
\PNA{The construction of the $\epsilon$-neighborhood is a major issue of the proposed methodology.} Since this operation could be extremely demanding, optimal
numerical implementation should be employed. In opposite with the Cell-linked
method proposed by Allen \& Tildesley \cite{Allen_et_al_1989} where
the definition of a numerical tolerance is required for auxiliar
algorithms like Crossing Number \cite{Shimrat_1962} or Correct
Even-Odd \cite{Galetzka_et_al_2017}, \PNA{the proposed methodology adopted herein is linked to the shape function neighbourhood. In order to avoid the aforementioned error} prone
techniques, we endeavour to exploit the meshfree benefits of
the \acrshort{lme} approximants to introduce a node-linked method. Due
to the global definition of this interpolation technique, the classical
interpretation of the \acrshort{fem} element typically adopted in
\acrshort{mpm} simulations is not required anymore. \DM{Intead, a} linked
list of tributary nodes is reconstructed with the closest node as
starting point at each time step \DM{deleted instead}. Therefore, each node is assigned to an
unique linked-list of those particles which are closest to it. This
approach allows to reconstruct easily a search list with the support
of the initial connectivity of the mesh and offers a suitable basis
for an easy implementation of other interpolation techniques with
minimum coding effort. A detailed explanation of the proposed
algorithm can be found in \ref{sec:node-linked-method-1}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Cases of study and discussion}
\label{sec:3}

\DM{The proposed approach to overcome the two main shortcomings of EighenMPM \cite{Zhang_EE_2020} has been evaluated with "two benchmark tests"¿?. } \DM{DELETE: As noticed in \cite{Zhang_EE_2020}, the EigenMPM have two important
shortcomings}. The first one deals with the presence of stress instabilities provoked by the spatial discretization, appearing even
when the \acrshort{gimp} shape functions are employed. To overcome it, \acrshort{lme}
approximants are introduced as an
alternative to the existing interpolation techniques. Its
performance in mitigating spurious stress oscillations under the \acrshort{mpm}~\cite{Wobbes2020,Molinos2020} framework helps to enhance
notoriously the quality of the results as it is observed in Section
\ref{sec:3.1}, where both interpolation techniques
are compared by carrying out an eigenerosion simulation. The second
limitation is concerned to its inability to simulate properly quasi-brittle
fracture \cite{Navas_2018_ES}. It can be solved through the
eigensoftening algorithm described in Section \ref{sec:2.3}. A proof
of it is exposed in Section \ref{sec:3.3}, where experimental results
of a drop-weight impact test are compared with experimental ones as well as eigensoftening computations. \PNA{FALTA LO DE BUI}.

\subsection{Edge-cracked square panel in mode I}
\label{sec:3.1}

The aim of the problem presented hereinafter is to assess the capability of the \acrshort{lme}
approximants to improve the result \textit{versus} the standard linear interpolation. The application consists of a square plate of size $H = 1$ containing
an initial edge crack of length $0.25 \cdot H$ loaded in a pure mode I
by displacement control on the outer flanks of the plate (See Figure
\ref{fig:geometry-cracked-panel-mode-I} for details).
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{Figure-Mode_I}
  \caption{Geometry and boundary condition of the drop-weight impact test.}
  \label{fig:geometry-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\PNA{UNIDADES}

@@

The constitutive model considered in the numerical experiment is a
linear-elastic Hookean material, whose Young's modulus \gls{E} = 1.06, the
Poisson's ratio \gls{nu} = 0.333, and critical energy-release rate
\gls{Gf} = 0.0001. This set of parameters has been previously employed by Pandolfi~\textit{et al.}~\cite{Pandolfi_2012}. In
\acrshort{mpm}, two different discretization are required. On the one hand,
a cartesian grid of nodes is considered with a nodal spacing value of
0.025. On the other hand, the plate will be modeled with a initial
layout of four particles per element and occupying the Gauss quadrature
positions.\\

Figure \ref{fig:Reactions-cracked-panel-mode-I} clearly shows the
presence of wiggles in the reaction-displacements curve, in both loading and post-failure stages,
when linear interpolation technique is used. In contrast, \acrshort{lme} simulation does not
exhibit these spurious oscillations and remains linear until the failure. About the strength,
linear interpolation produces stiffer results than
\acrshort{lme}. This can be attributed to imprecision in the stress
field due to grid-crossing phenomena.\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{Figure-Reactions-Mode-I}
  \caption{Evolution of the reaction forces plotted \textit{versus}
    the imposed displacement for linear interpolation technique and
    the \acrshort{lme} approximants.}
  \label{fig:Reactions-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Additionally, Figure \ref{fig:Stress-cracked-panel-mode-I}
exhibits more clearly the difference between linear interpolation and
\acrshort{lme}, since numerical instability becomes rather significant
when particles crosses the boundary of the element in the linear
case, which is observed at 60 seconds. \PNA{Although, apparently, the differences in the obtained peaks are not extremely significant, it can be owing to the aforementioned hookean material employed~\cite{Zhang_EE_2020}. When more sophisticated constitutive models will be employed, severe
%if hyperelastic materials are consider for future simulations,
inaccuracies may appear in the stress field that could affect dramatically to the
final result. Focusing on the post failure behavior, it can be seen how \acrshort{lme} produces soft stress
field evolution even after breaking.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\subfigure[t = 25]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0100-mode-I.png}}
\subfigure[t = 60]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0250-mode-I.png}}
\subfigure[t = 100]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0398-mode-I.png}}
\caption{Evolution of the stress tensor magnitude for linear
  interpolation (pictures in the left side), and \acrshort{lme}
  (pictures in the right side). Both simulations performed with an
  eigenerosion algorithm.}
\label{fig:Stress-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
An important consideration regarding the presence of oscillations once
both parts of the panel are separated is observed in Figure
\ref{fig:Reactions-cracked-panel-mode-I}. These phenomena should not be
attributed to the eigenerosion algorithm since the fracture process is
over. They are due to the dynamic nature of the solver: \PNA{once the energy is release dinamically by the crack, the waves propagate along the domain.}

\subsection{Drop-weight impact test}
\label{sec:3.2}
\PNA{This section is devoted to proof
the accuracy of the eigensoftening algorithm within the \acrshort{mpm} framework when the behavior
of quasi-brittle materials under dynamic loading is simulated. One of the most interesting examples of this
loading case is the three-point bending test on a concrete notched beam when it is conducted under impact loading. Indeed, this experimental test is appropriate to be reproduced numerically with the explicit solver proposed in this paper since
fracture occurs in a period of milliseconds and waves propagate fast along the high strength
materials involved in the aforementioned test.} Accordingly, small time steps are
required to achieve a numerically stable simulation.\\

@@

\begin{table}
  \centering
  \begin{tabular}[]{c c c c c c c}
    \hline
      &   $\rho$   & $\text{f}_c$ & $\text{f}_t$ & $G_F$ &   E   & $\text{d}_{max}$ \\
      & (kg/m$^3$) &     (MPa)    &     (MPa)    & (N/m) & (GPa) & (mm) \\
    \hline
Value &    2368    &     102.7    &      5.4     &  141  &   31  &  12 \\
    \hline
  \end{tabular}
  \caption[Mechanical properties of thje concrete]{Mechanical
    properties of the high strenght concrete.}
  \label{tab:concrete-properties}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

For the present research, a 2D setting of the experiment has been
adopted. Both the hammer and the concrete beam are explicitly
represented in order to properly capture the impact. Several levels of discretization were required to assess
the objectiveness of the obtained results. The results here exposed
are from a discretization of 30616 material points and 6297 nodes. To
achieve better results, an unestructured grid layout has been adopted
focusing the minimum nodal spacing, 0.47 mm, in the middle of the beam and in
the impact surface of the hammer. For this simulation, \gls{gamma-lme} was
fixed to 6. Once the accuracy of the \acrshort{lme} approximants have been proven
in Section \ref{sec:3.1}, this discretization has been adopted.
%without comparing
%with linear interpolation. Therefore \\

@@

\section{Node-linked method for $\epsilon$-neighborhood reconstruction}
\label{sec:node-linked-method-1}
\begin{algorithm}[H]
  \DontPrintSemicolon
  \KwInput{$I_0$ : Closest node for each particle}
  \KwInput{$I_{Nodes}$ : List of nodes close to each node}
  \KwInput{$Particles_I$ : List of particles close to each node}
  \KwOutput{$\text{neighborhood}_{\epsilon}$ : $\epsilon$-neighborhood for each particle}
  \For{$p$ to $N_p$}
  {
    For each $p$ get $I_0$\;
    Eval $I_{Nodes}(I_0)$ \;
    \For{$k \in I_{Nodes}(I_0)$}
    {
      $\text{neighborhood}_{\epsilon}^{k+1}$ = $\text{neighborhood}_{\epsilon}^{k} \cup Particles_I(k)$
    }
    \For{$p_{\epsilon} \in \text{neighborhood}_{\epsilon}$}
    {
      \If{$distance(p_{\epsilon},p) > \epsilon $}
      {
        pop $p_{\epsilon}$ from neighborhood$_{\epsilon}$\;
      }      
    }

>>>>

\newglossaryentry{nu}
{
  name=$\nu$,
  description={Poisson ratio}
}

\newglossaryentry{Ceps}
{
  name=$C_{\epsilon}$,
  description={Normalizing constant}
}

\newglossaryentry{heps}
{
  name=$h_{\epsilon}$,
  description={Band width}
}

\newglossaryentry{Wc}
{
  name=$w_{c}$,
  description={Critical opening displacement}
}

@@

\subsection{Fracture modelling approach}
\label{sec:2.3}
Within the context of \acrshort{mpm} formulation, fracture can be
modelled by failing particles according to a suitable failure
criterion. When material points are failed, they are assumed to have
null stress tensor. To reproduce this behaviour in the present
research, the eigensoftening algorithm is introduced in the 
\acrshort{mpm} framework as an alternative approach to the decohesion
model \cite{Zhenmao_2005,Schreyer_2002}. The eigensoftening concept
was originally developed by Navas {\it et al.}
(2017)\cite{Navas_2017_ES} as an extension for quasi-brittle materials
of the eigenerosion proposed by Pandolfi \& Ortiz
(2012)\cite{Pandolfi_2012} for fracture of brittle materials.\DM{The
  evaluation of both approaches \cite{Navas_2017_ES}}\MMP{against
  experimental measures}\DM{show that the eigenerosion algorithm significantly overestimates the tensile stress and the strain peaks, meanwhile eigensoftening captures the forces and crack
patterns accurately}. Furthermore,
the accuracy of this algorithm has been proven for complex fracture
patterns such as the ones obtained with fiber reinforces concrete (FRC),
\cite{Navas_2018_ES}.\\

The key idea behind the eigenerosion algorithm is the computation of the
energy-release rate attendant to the failure of material point $p$,
\begin{align}
  \label{eq:energy-release-EE}
&G_p^{k+1} = \frac{C_{\epsilon}}{m_p^{k+1}}  \sum_{x_q^{k+1} \in
  B_{\epsilon}(x_p^{k+1})} m_q W_q^{k+1}\\
  \label{eq:mass-EE}
&m_p^{k+1} =  \sum_{x_q^{k+1} \in
  B_{\epsilon}(x_p^{k+1})} m_q  
\end{align}
where $B_{\epsilon}(x_p^{k+1})$ is a $^n$dimensional sphere of radius $\epsilon$
centered at $x_p^{k+1}$. The particles which lie under this sphere are
known as the $\epsilon$-neighborhood of the material point (see
\cite{Pandolfi_2012}). This concept is conveniently sketched in
Figure \ref{fig:Failed-particles}.
%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.5\textwidth]{Figure-Particle-failed}
  \caption{Scheme of a linear cohesive law, where the shaded area is
    $G_f$, $f_t$ is the tensile strength, and $w_c$ is the critical
    opening displacement.}
  \label{fig:Failed-particles}
\end{figure}
%%%%%%%%%%%
Other parameters are the mass of the neighborhood $m_p^{k+1}$, the
current free-energy density per unit mass  $W_q^{k+1}$ and the
normalizing constant $C_{\epsilon}$, \PNA{which also defines the
  $\epsilon$-neighborhood size, being $\epsilon = C_\epsilon h$}, \MMP{ and
  $h$ is a measure of the distance between nodes.}

@@

reached. Once the crack, $w_n$, reaches the value of the critical
crack opening, $w_c$, a stress-free crack is attained.

@@

initialization was adopted. \MMP{Particularly the maximum principal stress
theory for brittle fracture was considered in \cite{Navas_2017_ES},
this allows to quantify the variation of the averaged strain energy density in the
$\epsilon$-neighborhood of the material point $\vec{x}_p^{k+1}$ as}

@@

where $m_p$ is the total mass of the
$\epsilon$-neighborhood,\DM{defined similarly to the eigenerosion
  method (Eq.~\eqref{eq:mass-EE}). The equivalent critical stress
  leads to a definition of the   averaged strain energy} in terms of the averaged

@@

\subsection{$\epsilon$-neighbourhood reconstruction : A node-linked method}
\label{sec:epsil-neighb-reconst}
\PNA{The construction of the $\epsilon$-neighbourhood is a major issue of the proposed methodology.} Since this operation could be extremely demanding, optimal
numerical implementation should be employed. In opposite with the Cell-linked
method proposed by Allen \& Tildesley \cite{Allen_et_al_1989} where
the definition of a numerical tolerance is required for auxiliar
algorithms like Crossing Number \cite{Shimrat_1962} or Correct
Even-Odd \cite{Galetzka_et_al_2017}, \PNA{the proposed methodology adopted herein is linked to the shape function neighbourhood. In order to avoid the aforementioned error} prone
techniques, we endeavour to exploit the meshfree benefits of
the \acrshort{lme} approximants to introduce a node-linked method. Due
to the global support of this interpolation technique, the classical
interpretation of the \acrshort{fem} element typically adopted in
\acrshort{mpm} simulations is not required anymore. Instead, a linked
list with the surrounding nodes for each node is defined at the initial
time. Also, each particle has to be associated with the
closest node instead with the belonging element as heretofore has been
done.
\MMP{Now lets define the concept of a particle \textit{tributary nodes} as the
list of nodes close to it among which the information transfers
occurs. Once the closest node are defined, the particle
posses automatically a list of \textit{tributary nodes}. This has two
immediate benefits, the first is that shape functions with global
support as \acrshort{lme} are easier to implement,
and the second is the increase of the efficiency of the $\epsilon$-neighbourhood
reconstruction. Regarding this last benefit, since the current
particle has a \textit{tributary nodes} list assigned to it, and
consequently each node has a linked list with a set of
\textit{tributary particles}\footnote{Notice that the list of the
  particles close to each node is unique for each node and is updated
  in each time step.} close to it, is possible to define efficiently a
search list with those particles close to the current particle.
In the end, particles out of the area of influence, Figure
\ref{fig:Failed-particles}, are pop out of the final
$\epsilon$-neighbourhood. A final remark about this new approach is
its ability to reconstruct easily a search list with the support of
the initial connectivity of the mesh. It provides a suitable basis for
an easy implementation of other interpolation techniques with minimum
coding effort. A detailed explanation of the proposed algorithm can be
found in \ref{sec:node-linked-method-1}.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Cases of study and discussion}
\label{sec:3}

\DM{The proposed approach to overcome the two main shortcomings of
  EighenMPM \cite{Zhang_EE_2020} has been evaluated with ``three
  benchmark tests''}.
The first one deals with the presence of stress
instabilities provoked by the spatial discretization, appearing even when the \acrshort{gimp} shape functions are employed. To overcome it, \acrshort{lme}
approximants are introduced as an alternative to the existing interpolation techniques. Its
performance in mitigating spurious stress oscillations under the \acrshort{mpm}~\cite{Wobbes2020,Molinos2020} framework helps to enhance
notoriously the quality of the results as it is observed in Section
\ref{sec:3.1}, where both interpolation techniques
are compared by carrying out an eigenerosion simulation. The second
limitation is concerned to its inability to simulate properly quasi-brittle
fracture \cite{Navas_2018_ES}. It can be solved through the
eigensoftening algorithm described in Section \ref{sec:2.3}.\MMP{A proof
of it is exposed in Sections \ref{sec:3.2} \ref{sec:3.3}, where experimental results
are compared with experimental ones as well as eigensoftening
computations. In the first case comparing with a semicircular bending
test and in the second case with a drop-weight impact test.}

\subsection{Edge-cracked square panel in mode I}
\label{sec:3.1}
The aim of the problem presented hereinafter is to assess the
capability of the \acrshort{lme} approximants to improve the result
\textit{versus} the standard linear interpolation. The application
consists of a square plate of size\MMP{$H = 1\ [L] $} containing an initial edge
crack of length $0.25 \cdot H$ loaded in a pure mode I by displacement
control on the outer flanks of the plate (See Figure
\ref{fig:geometry-cracked-panel-mode-I} for details). 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{Figure-Mode_I}
  \caption{Geometry and boundary condition of the drop-weight impact test.}
  \label{fig:geometry-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The constitutive model considered in the numerical experiment is a
linear-elastic Hookean material, whose Young's modulus \MMP{\gls{E} = 1.06
$[M] / ([L] \cdot [T]^2)$, the Poisson's ratio \gls{nu} = 0.333, and
critical energy-release rate \gls{Gf} = 0.0001 $[M] / \cdot
[T]^2$. The absence of units is due to the current simulation will not be validated against experimental results, where
scale factor is relevant, but will be against an analytical
solution provided by Pandolfi \& Ortiz \cite{Pandolfi_2012}.} In
\acrshort{mpm}, two different discretization are required. On the one hand, a cartesian grid of
nodes is considered with a nodal spacing value of 0.025. On the other
hand, the plate will be modeled with a initial layout of four
particles per element and occupying the Gauss quadrature positions.\\

@@

Figure \ref{fig:Reactions-cracked-panel-mode-I} clearly shows the
presence of wiggles in the reaction-displacements curve, in both
loading and post-failure stages, when linear interpolation technique
is used. In contrast, \acrshort{lme} simulation does not exhibit these
spurious oscillations and remains linear until the failure. About the
strength, linear interpolation produces stiffer results than
\acrshort{lme}. This can be attributed to imprecision in the stress
field due to grid-crossing phenomena.\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{Figure-Reactions-Mode-I}
  \caption{Evolution of the reaction forces plotted \textit{versus}
    the imposed displacement for linear interpolation technique and
    the \acrshort{lme} approximants.}
  \label{fig:Reactions-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Additionally, Figure \ref{fig:Stress-cracked-panel-mode-I}
exhibits more clearly the difference between linear interpolation and
\acrshort{lme}, since numerical instability becomes rather significant
when particles crosses the boundary of the element in the linear
case, which is observed at 60 seconds. \PNA{Although, apparently, the
  differences in the obtained peaks are not extremely significant, it
  can be owing to the aforementioned hookean material
  employed~\cite{Zhang_EE_2020}. When more sophisticated constitutive
  models will be employed, severe inaccuracies may appear in the
  stress field that could affect dramatically to the final 
  result. Focusing on the post failure behavior, it can be seen how
  \acrshort{lme} produces soft stress field evolution even after
  breaking.} 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\subfigure[t = 25]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0100-mode-I.png}}
\subfigure[t = 60]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0250-mode-I.png}}
\subfigure[t = 100]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0398-mode-I.png}}
\caption{Evolution of the stress tensor magnitude for linear
  interpolation (pictures in the left side), and \acrshort{lme}
  (pictures in the right side). Both simulations performed with an
  eigenerosion algorithm.}
\label{fig:Stress-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
An important consideration regarding the presence of oscillations once
both parts of the panel are separated is observed in Figure
\ref{fig:Reactions-cracked-panel-mode-I}. These phenomena should not be
attributed to the eigenerosion algorithm since the fracture process is
over. They are due to the dynamic nature of the solver: \PNA{once the
  energy is release dinamically by the crack, the waves propagate
  along the domain.}

\subsection{Semicircular bending test}
\label{sec:3.2}

The second application to be considered is two semicircular
bending tests on Johnstone, which have been carried out by Lim {\it et
  al.} \cite{LIM_1993} and numerically reproduced by Wang {\it et
  al.}\cite{Wang_2020} under \acrshort{sph} framework within an
implicit algorithm. The aim of the test is to examine the predictive
capability of the eigensoftening algorithm under different kind of
mixed-mode loading conditions. This capability was previously within a
\acrshort{otm} implementation \cite{Navas_2018_ES}.

The geometry and boundary conditions are sketched in Figure 
\ref{fig:geometry-Semicircular-bending-test}. The experiment consist in two
semicircular specimens with a radius of 47.5 mm and a thickness of
20 mm are supported by two rollers at a span of 47.5 mm and loaded by
another roller on its mid-span. Notches with a length of 16.6 mm are
created with different notch angles of 0$^º$ and 60$^º$ with respect
to the vertical axis to investigate the influences of the
notch angles on the peak load.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.6\textwidth]{./Figure-Semicircular-bending-test}
  \caption{Geometry and boundary condition of the semicircular bending
    test.}
  \label{fig:geometry-Semicircular-bending-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The material properties of Johnstone are listed in Table \ref{tab:Johnstone-properties}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}
  \centering
  \begin{tabular}[]{l l}
    \hline
    Young's modulus (\gls{E})   & 0.4\ $GPa$       \\
    Poisson ratio (\gls{nu})    & 0.25           \\
    Density (\gls{rho})         & 1.54\ $g/cm^3$ \\
    Tensile strength (\gls{ft}) & 0.6\ $MPa$       \\
    Critical opening displacement (\gls{Wc}) & 0.015\ $mm$ \\
    \hline
  \end{tabular}
  \caption[Mechanical properties of Johnstone. ]{Material properties of Johnstone.}
  \label{tab:Johnstone-properties}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In the tests, particles were uniformly distributed with an almost
constant separation of 1 mm between them, resulting in a total number of 4672
\acrshort{mpm} particles for generating the specimen. Additionally a
regular cartesian background grid of 1917 nodes with 1.5 mm of nodal distance
 was generated.

% The velocities of the top boundary particles are vx =
% 0,vy = − 2 mm/s, whereas the velocities of the bottom boundary
% particles are as vy = 0.

%     \gls{heps}
%     \gls{Ceps}

% As Johnstone rocks are classified as soft rocks with a relatively
% small elastic Young's modulus. Ramp loading is applied to alleviate the dynamic
% effect due to the sudden loading.

\subsection{Drop-weight impact test}
\label{sec:3.3}
\PNA{This section is devoted to proof
the accuracy of the eigensoftening algorithm within the \acrshort{mpm} framework when the behavior
of quasi-brittle materials under dynamic loading is simulated. One of the most interesting examples of this
loading case is the three-point bending test on a concrete notched beam when it is conducted under impact loading. Indeed, this experimental test is appropriate to be reproduced numerically with the explicit solver proposed in this paper since
fracture occurs in a period of milliseconds and waves propagate fast along the high strength
materials involved in the aforementioned test.} Accordingly, small time steps are
required to achieve a numerically stable simulation.\\

@@

\begin{table}
  \centering
  \begin{tabular}[]{c c c c c c c}
    \hline
      &   $\rho$   & $\text{f}_c$ & $\text{f}_t$ & $G_F$ &   E   & $\text{d}_{max}$ \\
      & (kg/m$^3$) &     (MPa)    &     (MPa)    & (N/m) & (GPa) & (mm) \\
    \hline
Value &    2368    &     102.7    &      5.4     &  141  &   31  &  12 \\
    \hline
  \end{tabular}
  \caption[Mechanical properties of thje concrete]{Mechanical
    properties of the high strenght concrete.}
  \label{tab:concrete-properties}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
For the present research, a 2D setting of the experiment has been
adopted. Both the hammer and the concrete beam are explicitly
represented in order to properly capture the impact. Several levels of
discretization were required to assess the objectiveness of the
obtained results. The results here exposed are from a discretization
of 30616 material points and 6297 nodes. To achieve better results, an
unestructured grid layout has been adopted focusing the minimum nodal
spacing, 0.47 mm, in the middle of the beam and in the impact surface
of the hammer. For this simulation, \gls{gamma-lme} was fixed to
6. Once the accuracy of the \acrshort{lme} approximants have been
proven in Section \ref{sec:3.1}, this discretization has been adopted.
%without comparing
%with linear interpolation. Therefore \\

@@

\section{Node-linked method for $\epsilon$-neighbourhood reconstruction}
\label{sec:node-linked-method-1}
\begin{algorithm}[H]
  \DontPrintSemicolon
  \KwInput{$I_0$ : Closest node for each particle}
  \KwInput{$I_{Nodes}$ : List of nodes close to each node}
  \KwInput{$Particles_I$ : List of particles close to each node}
  \KwOutput{$\text{neighbourhood}_{\epsilon}$ : $\epsilon$-neighbourhood for each particle}
  \For{$p$ to $N_p$}
  {
    For each $p$ get $I_0$\;
    Eval $I_{Nodes}(I_0)$ \;
    \For{$k \in I_{Nodes}(I_0)$}
    {
      $\text{neighbourhood}_{\epsilon}^{k+1}$ = $\text{neighbourhood}_{\epsilon}^{k} \cup Particles_I(k)$
    }
    \For{$p_{\epsilon} \in \text{neighbourhood}_{\epsilon}$}
    {
      \If{$distance(p_{\epsilon},p) > \epsilon $}
      {
        pop $p_{\epsilon}$ from neighborhood$_{\epsilon}$\;
      }      
    }
